\textsc{PAPERCLIP}: Associating Astronomical Observations and Natural Language with Multi-Modal Models

### 2024-03-05 paper/main.tex
Add blurbs about raw abs trained model in app

\subsection{Validation Metrics}

@@

@@

\SM{Describe science usefulness, see appendix.}

@@

\subsection{Distribution of Learned Representations}

@@

\section{Evaluation of Model Trained on Raw Abstracts}

@@

\begin{table}[h!]
  \centering
  \renewcommand{\arraystretch}{0.1}
  \begin{tabular}{m{3cm} m{3.2cm} m{3.2cm} m{5cm}}
      \toprule
      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ {\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ {\textcolor{deepblue}{(abstract fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_0.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_0.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_0.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_abs_0.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_1.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_1.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_1.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_abs_1.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_2.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_2.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_2.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_abs_2.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_3.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_3.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_3.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_abs_3.txt}} \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{Text associations from a curated list most closely matching a given image query, for both the \textcolor{deepblue}{abstract fine-tuned} and \textcolor{deeppurple}{base} (CLIP-ViT-B/16) models. The `ground truth' LLM-summarized abstract is shown in the right column.}
  \label{tab:itt_abs}
\end{table}

>>>>

\subsection{Quantitative Evaluation}

\paragraph*{Validation metrics during training}

@@

\paragraph*{Distribution of text-image cosine similarities}

\paragraph*{Retrieval accuracy}

@@

@@

Note that we choose to illustrate qualitative performance on text and image retrieval using the model fine-tuned on summarized abstracts, rather than raw abstracts. We show analogous results for the model fine-tuned on raw abstracts in App.~\ref{app:eval_raw}. Although the two models show very similar quantitative performance on retrieval metrics (as shown in Fig.~\ref{fig:sim_valtrain}), they exhibit characteristic behavior in terms of objects (images/text) retrieved. We emphasize that for scientific usefulness, retrieving the `most similar' objects would not be the goal, but rather identifying interesting candidates to follow up on; both models perform sensibly in this regard.

@@

\section{Evaluation of Model Trained on Raw Abstracts}
\label{app:eval_raw}

@@

\begin{table}[h!]
  \centering
  \renewcommand{\arraystretch}{0.1}
  \begin{tabular}{m{3cm} m{3.2cm} m{3.2cm} m{3.2cm}}
      \toprule
      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ {\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ {\textcolor{deepblue}{(abstract fine-tuned)}} & \centering  \textbf{Top-4 text} \\ {\textcolor{deepred}{(summary fine-tuned)}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_0.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_0.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_0.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_0.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_1.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_1.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_1.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_1.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_2.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_2.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_2.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_2.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_abs_3.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_3.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_abs_3.txt} &  {\scriptsize \verbatiminput{\thedatafolder/sci_itt_3.txt}} \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{Text associations from a curated list most closely matching a given image query (first column), for the \textcolor{deeppurple}{base} (CLIP-ViT-B/16) model, \textcolor{deepblue}{abstract fine-tuned} model, and \textcolor{deepred}{summary fine-tuned} model.}
  \label{tab:itt_abs}
\end{table}
